Arrows for Parallel Computations

### 2017-08-16 paper/content/utilityFunctions.tex
minor. added notes talk with oleg
moved trivial definitions of matrix multiplication to the appendix

\olcomment{more clear structure of the appendix? 1) Arrow utilities, 2) our things we omit for being too simple, 3) other's things we use.}
\section{Utility Arrows}\label{utilfns}
The 
|second| combinator: 
\begin{code}
second :: Arrow arr => arr a b -> arr (c, a) (c, b)
second f = arr swap >>> first f >>> arr swap
	where swap (x, y) = (y, x)
\end{code}
is a mirrored version of |first|, used, \eg in
definition of |***|.

@@

\section{Omitted Funtion Definitions}
\label{app:omitted}
We have omitted some function definitions in the main text for
brevity, and redeem this here. 
%
We warp Eden's build-in Futures in PArrows as in
Figure~\ref{fig:RDFuture}, where |rd| is the accessor function for the |RD| wrapped inside |RemoteData|. Furthermore, in order for these |Future| types to fit with the |ArrowParallel| instances we gave earlier, we have to give the necessary |NFData| and |Trans| instances, the latter are only needed in Eden. The |Trans| instance does not have any functions declared as the default implementation suffices here. Furthermore, because |MVar| already ships with a |NFData| instance, we only have to supply a simple delegating |NFData| instance for our |RemoteData| type, where |rd| simply unwraps |RD|. The |Trans| instance does not have any functions declared as the default implementation suffices:

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel arrow running on two instances of |f +++ g| as described above. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{Definition of parEval2.}
	\label{fig:parEval2}
\end{figure}

>>>>

\section{Utility Arrows}\label{utilfns}

\mbcomment{Text außenrum bauen, siehe Omitted Function Definitions}
The 
|second| combinator: 
\begin{code}
second :: Arrow arr => arr a b -> arr (c, a) (c, b)
second f = arr swap >>> first f >>> arr swap
	where swap (x, y) = (y, x)
\end{code}
is a mirrored version of |first|, used, \eg in
definition of |***|.

@@

\section{Omitted Function Definitions}
\label{app:omitted}
We have omitted some function definitions in the main text for
brevity, and redeem this here. 
%
We warp Eden's build-in Futures in PArrows as in
Figure~\ref{fig:RDFuture}, where |rd| is the accessor function for the |RD| wrapped inside |RemoteData|. Furthermore, in order for these |Future| types to fit with the |ArrowParallel| instances we gave earlier, we have to give the necessary |NFData| and |Trans| instances, the latter are only needed in Eden. The |Trans| instance does not have any functions declared as the default implementation suffices here. Furthermore, because |MVar| already ships with a |NFData| instance, we only have to supply a simple delegating |NFData| instance for our |RemoteData| type, where |rd| simply unwraps |RD|. The |Trans| instance does not have any functions declared as the default implementation suffices:

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel arrow running on two instances of |f +++ g| as described above. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{Definition of |parEval2|.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the ommited definitions of |prMMTr|, - which calculates $AB^T$ for two matrices $A$ and $B$, |splitMatrix| - which splits the a matrix into chunks, and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]

splitMatrix :: Int -> Matrix -> [[Matrix]]
splitMatrix size matrix = map (transpose . map (chunksOf size)) $ chunksOf size $ matrix

matAdd = chunksOf (dimX x) $ zipWith (+) (concat x) (concat y)
\end{code}
	\caption{Definition of |prMMTr|, |splitMatrix| and |matAdd|.}
	\label{fig:torus_example_rest}
\end{figure}
